### 2016-09-30 Supplement.tex
Initialise supporting information TEX file

\documentclass[12pt]{article}
\usepackage[top=1.25in, bottom=1in, left=1.25in, right=1in]{geometry}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{lineno}
\usepackage{datetime}
\setkomafont{\rmfamily\bfseries\boldmath}
\usepackage{wrapfig,floatrow}
\usepackage{float}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{fancyhdr}
\usepackage{titling}
\usepackage{bm}
\floatstyle{plain}
\restylefloat{figure}

\newcommand\mydef{\mathrel{\stackrel{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}

\def\wl{\par \vspace{\baselineskip}} %defines \wl as a line skip.

\renewcommand{\thesection}{S1.\arabic{section}}
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\renewcommand{\thepage}{S1-\arabic{page}}

\makeatletter %% With ams
\def\tagform@#1{\maketag@@@{(S1-\ignorespaces#1\unskip\@@italiccorr)}}
\makeatother

\makeatletter
\makeatletter \renewcommand{\fnum@figure}
{\figurename~S1-\thefigure}
\makeatother

\newcommand{\subtitle}[1]{%
  \posttitle{%
    \par\end{center}
    \begin{center}\large#1\end{center}
    \vskip0.5em}%
}

\setcitestyle{round,aysep={},yysep={;}}

\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
\begin{flushleft}
  \fontsize{18}{16}
  \textbf{\@title}
  \vspace{8 mm}
  \fontsize{12}{14}
  \@author
\end{flushleft}\egroup
}
\makeatother

%\newlength{\saveparindent} %Makes ragged right alignment for Evolution
\setlength{\saveparindent}{\parindent} %Does not suppress paragraph indents
%\raggedright
\setlength{\parindent}{\saveparindent}

\renewcommand\refname{LITERATURE CITED}
\newcommand*\mystrut[1]{\vrule width0pt height1pt depth#1\relax}
\newcommand={\mathrel{\stackrel{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}

\newenvironment{phv}{\fontfamily{phv}\selectfont}{\phv}

\title{Evolution of pre-copulatory and post-copulatory inbreeding behaviour \\ \vspace{10 mm} {\sffamily\Large\bfseries Supplemental results showing change in allele values over different parameter combinations, and allele values over generations for single replicates}}

\author{\\ \textbf{A. Bradley Duthie\textsuperscript{1,2} and Jane M. Reid\textsuperscript{1}}\\ \textit{\textsuperscript{1} Institute of Biological and Environmental Sciences, School of Biological Sciences, Zoology Building, Tillydrone Avenue, University of Aberdeen, Aberdeen AB24 2TZ, United Kingdom \\ \hspace{10 mm} \textsuperscript{2} E-mail: aduthie@abdn.ac.uk}}

\date{}

\linenumbers
\doublespacing

\pagestyle{fancy}
\lhead{DUTHIE ET AL}
\rhead{PRE- & POST-COPULATORY (SUPPLEMENT)}
\renewcommand{\headrulewidth}{0pt}

\begin{document}

\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
                                   {-3.5ex \@plus -1ex \@minus -.2ex}%
                                   {2.3ex \@plus.2ex}%
                                   {\sffamily\Large\itshape}}% from \Large
\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\sffamily\Large\bfseries}}% from \large
\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\sffmaily\large\bfseries}}% from \normalsize
\makeatother

\begin{phv}
\thispagestyle{empty} % single page disable page number

\maketitle

\vspace{10 mm}

\noindent \textbf{Here we present supplemental results from an individual-based model in which alleles underlying polyandry (\bm{$P_{a}$}), pre-copulatory inbreeding strategy (\bm{$M_{a}$}) and post-copulatory inbreeding strategy (\bm{$F_{a}$}) affect phenotypic values for .}

\clearpage

\vspace*{5 mm}

{\Large Table of Contents}

\hrulefill

\contentsline {section}{\tocsection {}{}{\small Figure S\ref{allcombs}: All cost combinations}}{S3}
\contentsline {section}{\tocsection {}{}{\small Figure S\ref{extra_MfPf}: All cost combinations}}{S4}
\contentsline {section}{\tocsection {}{}{\small Figure S\ref{SI_M00_P00_E00}: All cost combinations}}{S5}
\contentsline {section}{\tocsection {}{}{\small Figure S\ref{SI_M02_P00_E00}: All cost combinations}}{S6}
\contentsline {section}{\tocsection {}{}{\small Figure S\ref{SI_M00_P02_E00}: All cost combinations}}{S7}
\contentsline {section}{\tocsection {}{}{\small Figure S\ref{SI_M00_P00_E02}: All cost combinations}}{S8}
\contentsline {section}{\tocsection {}{}{\small Figure S\ref{SI_M02_P02_E00}: All cost combinations}}{S9}
\contentsline {section}{\tocsection {}{}{\small Figure S\ref{SI_M02_P00_E02}: All cost combinations}}{S10}
\contentsline {section}{\tocsection {}{}{\small Figure S\ref{SI_M00_P02_E02}: All cost combinations}}{S11}
\contentsline {section}{\tocsection {}{}{\small Figure S\ref{SI_M02_P02_E02}: All cost combinations}}{S12}

\hrulefill

\end{phv}

\clearpage

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.85]{figures/evo_combs.pdf}
\end{center}
\caption{Some new costs}
\label{allcombs}
\end{figure}

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.8]{figures/extra_MfPf.pdf}
\end{center}
\caption{A key caveat to the results}
\label{extra_MfPf}
\end{figure}

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.8]{figures/SI_M00_P00_E00.pdf}
\end{center}
\caption{A key caveat to the results}
\label{SI_M00_P00_E00}
\end{figure}

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.8]{figures/SI_M02_P00_E00.pdf}
\end{center}
\caption{A key caveat to the results}
\label{SI_M02_P00_E00}
\end{figure}

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.8]{figures/SI_M00_P02_E00.pdf}
\end{center}
\caption{A key caveat to the results}
\label{SI_M00_P02_E00}
\end{figure}

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.8]{figures/SI_M00_P00_E02.pdf}
\end{center}
\caption{A key caveat to the results}
\label{SI_M00_P00_E02}
\end{figure}

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.8]{figures/SI_M02_P02_E00.pdf}
\end{center}
\caption{A key caveat to the results}
\label{SI_M02_P02_E00}
\end{figure}

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.8]{figures/SI_M02_P00_E02.pdf}
\end{center}
\caption{A key caveat to the results}
\label{SI_M02_P00_E02}
\end{figure}

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.8]{figures/SI_M00_P02_E02.pdf}
\end{center}
\caption{A key caveat to the results}
\label{SI_M00_P02_E02}
\end{figure}

\clearpage
\begin{figure}
\begin{center}				
\includegraphics[scale=0.8]{figures/SI_M02_P02_E02.pdf}
\end{center}
\caption{A key caveat to the results}
\label{SI_M02_P02_E02}
\end{figure}

\end{document}