### 2018-02-26 presentation/inbreeding.Rmd
Sketch out some introductory slides

---
title: "When, why, and how to avoid, prefer, or tolerate inbreeding"
subtitle: "https://bradduthie.shinyapps.io/inbreeding"
author: "Brad Duthie"
date: "6 March 2018"
output: 
  ioslides_presentation: 
    widescreen: yes
runtime: shiny
---

<style>
div.footnotes {
  position: absolute;
  bottom: 0;
  margin-bottom: 10px;
  width: 80%;
  font-size: 0.6em;
}
</style>

<script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script>

<script>
  $(document).ready(function() {
    $('slide:not(.backdrop):not(.title-slide)').append('<div class=\"footnotes\">');

    $('footnote').each(function(index) {
      var text  = $(this).html();
      var fnNum = (index+1).toString().sup();
      $(this).html(text + fnNum);

      var footnote   = fnNum + ': ' + $(this).attr('content') + '<br/>';
      var oldContent = $(this).parents('slide').children('div.footnotes').html();
      var newContent = oldContent + footnote;
      $(this).parents('slide').children('div.footnotes').html(newContent);
    });
  });
</script>

```{r, echo = FALSE}
initialise_inds <- function(N = 200){
    ID      <- matrix(data = 0, nrow = N, ncol = 30);
    ID[,1]  <- 1:N;
    ID[,2]  <- rbinom(n = N, size = 1, prob = 0.5);
    ID[,3]  <- sample(x = 1:10, size = N, replace = TRUE);
    ID[,4]  <- sample(x = 1:10, size = N, replace = TRUE);
    ID[,5]  <- rep(x = 0, times = N);
    ID[,6]  <- rep(x = -1, times = N);
    ID[,7]  <- rep(x = -1, times = N);
    ID[,8]  <- rep(x = 0, times = N);
    ID[,9]  <- rep(x = -1, times = N);
    ID[,10] <- rep(x = 0, times = N);
    return(ID);
}

initialise_Rmat <- function(ID){
    rows <- dim(ID)[1];
    Rmat <- matrix(data = 0, nrow = rows, ncol = rows);
    diag(Rmat) <- 1;
    Rmat <- cbind(ID[,1], Rmat);
    return(Rmat);
}

Rbuild <- function(Rmat, ID, Rmat_new){
    inds <- dim(ID)[1];
    for(i in 1:inds){
        if(ID[i, 6] %in% Rmat[,1] == FALSE){
            MomP <- -1;
        }else{
            MomP <- which(Rmat[,1] == ID[i, 6]);
        }
        if(ID[i, 7] %in% Rmat[,1] == FALSE){
            DadP <- -1;
        }else{
            DadP <- which(Rmat[,1] == ID[i, 7]);
        }
        # Inbreeding coefficient
        if(MomP < 0 | DadP < 0){
            Rmat_new[i, i+1] <- 1; 
        }else{
            Rmat_new[i, i+1] <- 1 + Rmat[MomP, DadP+1]; 
        }
        # Kinship coefficient
        j <- 1;
        while(j < i){
            if(ID[j, 6] %in% Rmat[,1] == FALSE){
                Om <- -1;
            }else{
                Om <- which(Rmat[,1] == ID[j, 6]);
            }
            if(ID[j, 7] %in% Rmat[,1] == FALSE){
                Od <- -1;
            }else{
                Od <- which(Rmat[,1] == ID[j, 7]);
            }

            if(ID[i,5] == 0 & ID[j,5] == 0){
                if(Om < 0 | MomP < 0){
                    w <- 0;
                }else{
                    if(MomP == Om){
                        w <- 0.5 * Rmat[MomP, Om + 1];
                    }else{
                        w <- Rmat[MomP, Om + 1];
                    }
                }
                if(Od < 0 | MomP < 0){
                    y <- 0;
                }else{
                    y <- Rmat[MomP, Od + 1];
                }
                if(Om < 0 | DadP < 0){
                    x <- 0;
                }else{
                    x <- Rmat[DadP, Om + 1];
                }
                if(Od < 0 | DadP < 0){
                    z <- 0;
                }else{
                    if(DadP == Od){
                        z <- 0.5 * Rmat[DadP, Od + 1];
                    }else{
                        z <- Rmat[DadP, Od + 1];
                    }
                }
            }
            rMmuP <- 0.5 * (w + y);
            rDmuP <- 0.5 * (x + z);
            rval  <- 0.5 * (rMmuP + rDmuP);

            Rmat_new[i, j+1] <- rval;
            Rmat_new[j, i+1] <- rval;
            j <- j + 1;
        }
    }

    for(i in 1:inds){
        for(j in 1:inds){
            if(ID[i, 6] == ID[j, 1]){
                if(ID[j, 6] == -1){
                    Rmat_new[i, j+1] = 0.25;
                    Rmat_new[j, i+1] = 0.25;
                }    
            }
            if(ID[i, 7] == ID[j, 1]){
                if(ID[j, 7] == -1){
                    Rmat_new[i, j+1] = 0.25;
                    Rmat_new[j, i+1] = 0.25;
                }    
            }
        }
    }

    return(Rmat_new);
}

mateselect <- function(ID, Rmat_new){
    fem <- ID[ID[,2] == 0 & ID[,5] == 0,];
    mal <- ID[ID[,2] == 1 & ID[,5] == 0,];
    for(i in 1:dim(fem)[1]){
        strat <- sum(fem[i,11:20]);
        vals  <- rep(0, dim(mal)[1]);
        fpos  <- which(Rmat_new[,1] == i); 
        for(j in 1:length(vals)){
            mpos    <- which(Rmat_new[,1] == j);
            kval    <- Rmat_new[i, j+1];
            vals[j] <- kval * strat;
        }
        if(strat < 0){
            weight <- 1 / (1 + (-1 * vals));
            probs  <- weight / sum(weight);
        }else{
            weight <- 1 + vals;
            probs  <- weight / sum(weight);
        }
        mate    <- sample(x = 1:dim(mal)[1], size = 1, pr = probs);
        mateID  <- mal[mate,1];
        matepos <- which(ID[,1] == mateID);
        ownID   <- fem[i, 1];
        ownpos  <- which(ID[,1] == ownID);
        ID[ownpos, 9]  <- mateID;
        ID[matepos, 9] <- ownID;
    }
    return(ID);
}

make_offspring <- function(ID, Rmat_new, nn = 6){
    start_ID <- max(ID[,1]) + 1;
    for(i in 1:dim(ID)[1]){
        if(ID[i, 2] == 0 & ID[i, 9] > 0 & ID[i, 5] == 0){
            mum    <- ID[i,];
            dadpos <- which(ID[,1] == ID[i, 9]);
            dad    <- ID[dadpos,];
            offs   <- matrix(data = 0, nrow = nn, ncol = dim(ID)[2]);
            offs[,1] <- start_ID:(start_ID+nn-1);
            start_ID <- start_ID + nn;
            offs[,2] <- rbinom(n = nn, size = 1, prob = 0.5);
            offs[,3] <- mum[3];
            offs[,4] <- mum[4];
            offs[,6] <- mum[1];
            offs[,7] <- dad[1];
            mrpos    <- which(Rmat_new[,1] == mum[1]);
            drpos    <- which(Rmat_new[,1] == dad[1]);
            offs[,8] <- Rmat_new[mrpos, drpos+1];
            offs[,11:30] <- mum[11:30];
            dum_vec  <- rbinom(n = length(offs[,11:30]), size = 1, prob = 0.5);
            dummy    <- matrix(data = dum_vec, nrow = dim(offs)[1]);
            dummy[dummy > 0] <- NA;

        }
    }
}

produce_offs <- function(ID, Rmat_new, nn){
    # NOTE: This toy model has unrealistic recombination between parents!
    mums  <- sum(ID[,2] == 0 & ID[, 5] == 0 & ID[, 9] != 0);
    offs  <- matrix(data = 0, nrow = mums * nn, ncol = dim(ID)[2]);
    start <- 1;
    for(i in 1:dim(ID)[1]){  # Unrealistic recombination saves some time.
        if(ID[i,2] == 0 & ID[i, 5] == 0 & ID[i, 9] != 0){
            her_off     <- matrix(data = ID[i,], nrow = nn, ncol = dim(ID)[2], 
                                  byrow = TRUE);
            her_off[,2] <- rbinom(n = nn, size = 1, prob = 0.5);
            her_off[,6] <- ID[i,1];
            her_off[,7] <- ID[i,9];
            mum_p       <- which(Rmat_new[,1] == ID[i,1]);
            dad_p       <- which(Rmat_new[,1] == ID[i,9]);
            her_off[,8] <- Rmat_new[mum_p, dad_p + 1];
            her_off[,9] <- -1;
            dad_o       <- which(ID[,1] == ID[i,9]);
            dad_overlay <- matrix(data = ID[dad_o,], nrow = nn, 
                                  ncol = dim(ID)[2], byrow = TRUE);
            to_vec      <- rbinom(n = nn*dim(ID)[2], size = 1, prob = 0.5);
            to_overlay  <- matrix(data = to_vec, nrow = nn);
            to_overlay[,1:10] <- 1;
            her_off[to_overlay < 1] <- NA;
            her_off[is.na(her_off)] <- dad_overlay[is.na(her_off)];
            offs[start:(start+(nn-1)), 1:dim(ID)[2]] <- her_off;
            start <- start + nn;
        }
    }
    max_ID   <- max(ID[,1]);
    offs[,1] <- (max_ID + 1):(max_ID + dim(offs)[1]); 
    return(offs);
}

mortality <- function(offs, beta){
    depression        <- exp(-1 * beta * offs[,8]);
    alive             <- rbinom(n = dim(offs)[1], size = 1, prob = depression);
    offs[alive < 1,5] <- -1;
    offs              <- offs[offs[,5] > -1,];
    return(offs);
}

mutation  <- function(offs, mu){
    mutations <- rbinom(n = 20*dim(offs)[1], size = 1, prob = mu);
    tot_mut   <- sum(mutations);
    mut_mat   <- matrix(data = mutations, nrow = dim(offs)[1]);
    mu_eff    <- rnorm(n = tot_mut, mean = 0, sd = 2);
    mut_mat[mut_mat > 0] <- mu_eff;
    offs[,11:30] <- offs[,11:30] + mut_mat;
    return(offs);
}

retain <- function(ID, offs){
    retainers <- rep(x = 0, times = dim(ID)[1]);
    for(i in 1:dim(ID)[1]){
        ID_num <- ID[i, 1];
        ismum  <- sum(offs[,6] == ID_num);
        isdad  <- sum(offs[,7] == ID_num);
        ispar  <- ismum + isdad;
        if(ispar < 1){
            retainers[i] <- 1;
        }
    }
    to_remove <- which(retainers == 1);
    ID <- ID[-to_remove,];
    return(ID);
}

get_stats <- function(ID){
    mean_inbr <- mean(ID[,8]);
    mean_aval <- mean(ID[,11:20]);
    sd_aval   <- sd(ID[,11:20]);
    mean_nval <- mean(ID[,21:30]);
    sd_nval   <- sd(ID[,21:30]);
    stat_vec  <- c(mean_inbr, mean_aval, sd_aval, mean_nval, sd_nval);
    return(stat_vec);
}

immigration <- function(ID, imm, IDst){
    max_ID     <- max(ID[,1]);
    immigrants <- initialise_inds(N = imm);
    avalues    <- rnorm(n = 10 * imm, mean = IDst[2], sd = IDst[3]);
    nvalues    <- rnorm(n = 10 * imm, mean = IDst[4], sd = IDst[5]);
    a_mat      <- matrix(data = avalues, nrow = imm);
    n_mat      <- matrix(data = nvalues, nrow = imm);
    immigrants[,2]     <- 1;
    immigrants[,11:20] <- a_mat;
    immigrants[,21:30] <- n_mat;
    imm_IDs            <- (max_ID + 1):(max_ID + imm);
    immigrants[,1]     <- imm_IDs;
    ID <- rbind(ID, immigrants);
    return(ID);
}

new_gen <- function(ID, Rmat, cost = 0, imm = 5, beta = 1, 
                    mu = 0.01, Kf = 100, Km = 100, nn = 6){
    N       <- dim(ID)[1];
    death   <- rep(x = 0, times = N);
    females <- sum(ID[,2] == 0 & ID[,5] <= 1);
    males   <- sum(ID[,2] == 1 & ID[,5] <= 1);
    strat   <- apply(X = ID[,11:20], MARGIN = 1, FUN = sum);
    strat_c <- abs(strat) * cost;
    strat_d <- rbinom(n = N, size = 1, prob = strat_c);
    death   <- death + strat_d;
    death   <- death + (ID[,5] > 1);
    ID[,5]  <- -1 * death;
    ID      <- ID[order(ID[,2], decreasing = FALSE),];
    ID      <- ID[order(ID[,5], decreasing = FALSE),];
    livef   <- sum(ID[,2] == 0 & ID[,5] < 2 & ID[,5] > -1);
    livem   <- sum(ID[,2] == 1 & ID[,5] < 2 & ID[,5] > -1);
    if(livef > Kf){
        rm   <- livef - Kf;
        rows <- which(ID[,2] == 0 & ID[,5] < 2 & ID[,5] > -1);
        ID[rows[1:rm],5] <- -1;
    }
    if(livem > Km){
        rm   <- livem - Km;
        rows <- which(ID[,2] == 1 & ID[,5] < 2 & ID[,5] > -1);
        ID[rows[1:rm],5] <- -1;
    }
    ID <- rbind(ID[ID[,5] >= 0,], ID[ID[,5] < 0,]);

    Rmat_new <- initialise_Rmat(ID);
    Rmat_new <- Rbuild(Rmat, ID, Rmat_new);

    ID   <- mateselect(ID, Rmat_new);
    offs <- produce_offs(ID, Rmat_new, nn);
    offs <- mortality(offs, beta);
    offs <- mutation(offs, mu);
    IDst <- get_stats(ID);

    lgn  <- ID[ID[,5] == 0,];

    ID[ID[,5] >= 0, 5] <- -1;

    ID <- retain(ID, offs);
    ID <- rbind(ID, offs);

    ID <- immigration(ID, imm, IDst);

    results <- list(ID = ID, Rmat = Rmat_new, stats = IDst, last_gen = lgn);
    return(results);
}

mate_connect <- function(ID){
    living     <- ID[ID[,5] == 0,];
    fem_count  <- sum(living[,2] == 0);
    mal_count  <- sum(living[,2] == 1);
    livsor     <- living[order(living[,2]),];
    livsor[,3] <- 1:dim(livsor)[1];
    pointpairs <- matrix(data = 0, nrow = dim(livsor)[1], ncol = 8);
    pointpairs[,1] <- livsor[,1];
    for(i in 1:dim(livsor)[1]){
        if(livsor[i, 2] == 0){
            femID <- livsor[i, 1];
            malID <- livsor[i, 9];
            femps <- i;
            malps <- which(livsor[,1] == malID);
            pointpairs[i, 2] <- malID;
        }
    }
    pointpairs[,3] <- 1:dim(pointpairs)[1];
    distr <- seq(from = 0, to = 2*pi, length.out = dim(pointpairs)[1]);
    for(i in 1:dim(pointpairs)[1]){
        xplace            <- 2 * cos(distr[pointpairs[i, 3]]);
        yplace            <- 2 * sin(distr[pointpairs[i, 3]]);
        pointpairs[i, 5]  <- xplace;
        pointpairs[i, 6]  <- yplace;
    }
    for(i in 1:dim(pointpairs)[1]){
        if(i <= fem_count){
            matepos           <- which(pointpairs[,1] == pointpairs[i, 2]);
            pointpairs[i, 4]  <- matepos;
            pointpairs[i, 7]  <- pointpairs[matepos, 5];
            pointpairs[i, 8]  <- pointpairs[matepos, 6];
        }
    }
    return(pointpairs);
}
```

## Inbreeding

- Define relatedness

## Inbreeding depression

- What inbreeding depression is, exactly.

## Inbreeding depression

- What are the consequences of inbreeding depression.

## Inbreeding behaviour

- What do we exect to see given inbreeding depression?

- inbreeding avoidance
- inbreeding preference
- inbreeding tolerance

## Inbreeding theory

Describe Parker's model

- Resource and mate limited sexes
- Mate encounter, opportunity to outbreed assumed

## Inbreeding theory

<div class = "columns-2">

```{r echo = FALSE, fig.height = 4, fig.width = 5}
inputPanel(
  sliderInput("rr", label = "Relatedness", width="100%",
              min = 0.0, max = 1, value = 0, step = 0.05), 
  sliderInput("dd", label = "Inbreeding Depression", width="100%",
              min = 0.0, max = 1, value = 0, step = 0.1)
)
par(bg = 'transparent', mar = c(0.5, 0.5, 0.5, 0.5));
plot(x = 1:10, y = 1:10, xaxt = "n", yaxt = "n", xlab = "", ylab = "");
# Add see-saw function here to animate the balance
```

<br><br>

- Parker's model
- Add a see-saw plot that weighs relatedness versus inbreeding depression

$$\frac{n}{2} (1 + r)(1 - \delta) > \frac{n}{2}$$

Is inclusive fitness from inbreeding (left) greater than inclusive fitness from outbreeding?

$$\delta < \frac{r}{1+r}$$

</div>

## Interpreting Parker's model

Some key takeaways:

- Inbreeding avoidance is predicted given a sufficiently high threshold of inbreeding depression
- Inbreeding depression thresholds differ between females and males causing sexual conflict
- Females should prefer inbreeding below their inbreeding depression threshold (resource limited) <footnote content="The first awesoem footnote!">here</footnote>:
- Males should *tolerate* inbreeding below their threshold, but *never prefer* (sis will be fine)

##

```{r lmSim, fig.show='animate', echo = FALSE}
inputPanel(
  sliderInput("Imm", label = "Immigrants", width="100%",
              min = 0.0, max = 10, value = 0, step = 1), 
  sliderInput("ID", label = "Inbreeding Depression", width="100%",
              min = 0.0, max = 10, value = 0, step = 0.5), 
  sliderInput("cost", label = "Strategy Cost", width="100%",
              min = 0.0, max = 0.1, value = 0, step = 0.01), 
  sliderInput("mu", label = "Mutation Rate", width="100%",
              min = 0.0, max = 0.1, value = 0, step = 0.01), 
  actionButton("mate", "New Generation"),
  actionButton("reset", "Reset")
)

gen     <- reactiveValues();
iter    <- reactiveValues();
mtt     <- reactiveValues();

gen$dat  <- list(ID = NULL, Rmat = NULL, lgn = NULL);
iter$all <- 0;
iter$nue <- 0;
iter$kva <- 0;
iter$mkv <- 0;
mtt$dat  <- NULL;

Ds    <- observeEvent(input$mate, {
        if( is.null(gen$dat$ID) == TRUE | is.null(gen$dat$ID) == TRUE){
            gen$dat$ID   <- initialise_inds(N = 80);
            gen$dat$Rmat <- initialise_Rmat(gen$dat$ID);
        }else{
            gen$dat  <- new_gen(ID = gen$dat$ID, Rmat = gen$dat$Rmat, 
                                Kf = 40, Km = 40, beta = input$ID, 
                                imm = input$Imm, cost = input$cost, 
                                mu = input$mu);
            iter$all <- c(iter$all, mean(gen$dat$ID[,11:20]));
            iter$nue <- c(iter$nue, mean(gen$dat$ID[,21:30]));
            kvalues  <- gen$dat$Rmat;
            kvalues  <- kvalues[,-1];
            if(dim(kvalues)[1] > 80){
                kvalues <- kvalues[1:80, 1:80];
            }
            diag(kvalues) <- NA ;#diag(kvalues) * 0.5;
            iter$kva <- kvalues;
            iter$mkv <- c(iter$mkv, mean(kvalues, na.rm = TRUE));
            mtID     <- gen$dat$last_gen;
            mtt$dat  <- mate_connect(mtID);
        }
})

rs    <- observeEvent(input$reset, {
       gen$dat <- list(ID = NULL, Rmat = NULL);
       iter$all <- 0;
       iter$nue <- 0;
       iter$kva <- 0;
       iter$mkv <- 0;
})

renderPlot({
  par(bg = 'transparent');
  malle <- iter$all;
  mneut <- iter$nue
  mkval <- iter$mkv
  gens  <- 1:length(iter$all);
  if(length(gens) < 6){
      par(mfrow = c(1, 2));
      if(length(gens) > 1){
          dat <- mtt$dat;
          femrow <- 1:sum(dat[,2] > 0);
          malrow <- (sum(dat[,2] > 0) + 1):dim(dat)[1];
          plot(x = dat[, 5], y = dat[, 6], pch = 15, col = "blue", cex = 1.25,
               xlim = c(-2, 2), ylim = c(-2, 2), xaxt = "n", yaxt = "n", xlab = "",
               ylab = "", type = "n", bty = "n");
          text(x = -2, y = -2, pos = 4, col = "red", labels = "Males", cex = 1.75);
          text(x = -2, y =  2, pos = 4, col = "blue", labels = "Females", cex = 1.75);
          arrows(x0 = dat[femrow,5], x1 = dat[femrow,7], y0 = dat[femrow,6], 
              y1 = dat[femrow,8], length = 0, lwd = 2);
          points(x = dat[femrow, 5], y = dat[femrow, 6], pch = 15, col = "blue", cex = 1.25);
          points(x = dat[malrow, 5], y = dat[malrow, 6], pch = 16, col = "red", cex = 1.25);
          plot(x = 0:5, y = 0:5, type = "n", ylim = c(0, 1), cex.axis = 1.5,
               cex.lab = 1.5, xlab = "Generation", ylab = "Mean Kinship", 
               bg = "transparent");
          points(x = gens, y = mkval, pch = 20, cex = 1.5, type = "b",
                 lwd = 2, col = "black");
      }else{
          plot.new();
      }
  }else{
      par(mfrow = c(1, 2));
      dat <- mtt$dat;
      femrow <- 1:sum(dat[,2] > 0);
      malrow <- (sum(dat[,2] > 0) + 1):dim(dat)[1];
      plot(x = dat[, 5], y = dat[, 6], pch = 15, col = "blue", cex = 1.25,
           xlim = c(-2, 2), ylim = c(-2, 2), xaxt = "n", yaxt = "n", xlab = "",
           ylab = "", type = "n", bty = "n");
           arrows(x0 = dat[femrow,5], x1 = dat[femrow,7], y0 = dat[femrow,6], 
           y1 = dat[femrow,8], length = 0, lwd = 2);
      text(x = -2, y = -2, pos = 4, col = "red", labels = "Males", cex = 1.75);
      text(x = -2, y =  2, pos = 4, col = "blue", labels = "Females", cex = 1.75);
      points(x = dat[femrow, 5], y = dat[femrow, 6], pch = 15, col = "blue", cex = 1.25);
      points(x = dat[malrow, 5], y = dat[malrow, 6], pch = 15, col = "red", cex = 1.25);
      x0 <- length(gens) - 5;
      x1 <- length(gens) + 0;
      plot(x = x0:x1, y = 0:5, type = "n", ylim = c(0, 1), cex.axis = 1.5,
           cex.lab = 1.5, xlab = "Generation", ylab = "Mean Kinship", 
           bg = "transparent");
      points(x = gens, y = mkval, pch = 20, cex = 1.5, type = "b",
                 lwd = 2, col = "black");
  }
})
```

##

```{r lmSim3, fig.show='animate', echo = FALSE}
inputPanel(
  sliderInput("Imm3", label = "Immigrants", width="100%",
              min = 0.0, max = 10, value = 0, step = 1), 
  sliderInput("ID3", label = "Inbreeding Depression", width="100%",
              min = 0.0, max = 10, value = 0, step = 0.5), 
  sliderInput("cost3", label = "Strategy Cost", width="100%",
              min = 0.0, max = 0.1, value = 0, step = 0.01), 
  sliderInput("mu3", label = "Mutation Rate", width="100%",
              min = 0.0, max = 0.1, value = 0, step = 0.01), 
  actionButton("mate3", "New Generation"),
  actionButton("reset3", "Reset")
)

gen     <- reactiveValues();
iter    <- reactiveValues();

gen$dat  <- list(ID = NULL, Rmat = NULL);
iter$all <- 0;
iter$nue <- 0;
iter$kva <- 0;
iter$mkv <- 0;

Ds    <- observeEvent(input$mate3, {
        if( is.null(gen$dat$ID) == TRUE | is.null(gen$dat$ID) == TRUE){
            gen$dat$ID   <- initialise_inds(N = 80);
            gen$dat$Rmat <- initialise_Rmat(gen$dat$ID);
        }else{
            gen$dat  <- new_gen(ID = gen$dat$ID, Rmat = gen$dat$Rmat, 
                                Kf = 40, Km = 40, beta = input$ID3, 
                                imm = input$Imm3, cost = input$cost3, 
                                mu = input$mu3);
            iter$all <- c(iter$all, mean(gen$dat$ID[,11:20]));
            iter$nue <- c(iter$nue, mean(gen$dat$ID[,21:30]));
            kvalues  <- gen$dat$Rmat;
            kvalues  <- kvalues[,-1];
            if(dim(kvalues)[1] > 80){
                kvalues <- kvalues[1:80, 1:80];
            }
            diag(kvalues) <- NA ;#diag(kvalues) * 0.5;
            iter$kva <- kvalues;
            iter$mkv <- c(iter$mkv, mean(kvalues, na.rm = TRUE));
        }
})

rs    <- observeEvent(input$reset3, {
       gen$dat <- list(ID = NULL, Rmat = NULL);
       iter$all <- 0;
       iter$nue <- 0;
       iter$kva <- 0;
       iter$mkv <- 0;
})

renderPlot({
  par(bg = 'transparent');
  malle <- iter$all;
  mneut <- iter$nue
  mkval <- iter$mkv
  gens  <- 1:length(iter$all);
  if(length(gens) < 6){
      par(mfrow = c(1, 2));
      if(length(gens) > 0){
          par(mar = c(7, 5, 5, 5));
          hist(iter$kva, xlim = c(0, 1), ylim = c(0, 20), freq = FALSE, col = "black",
               yaxt = "n", ylab = "", main = "", xlab = "", 
               breaks = seq(from = 0, to = 1, by = 0.05), cex.axis = 1.5);
          title(xlab = "Kinship (k)", line = 4, cex.axis = 1.5, cex.lab = 1.5);
          box(lwd = 3);
          plot(x = 0:5, y = 0:5, type = "n", ylim = c(0, 1), cex.axis = 1.5,
               cex.lab = 1.5, xlab = "Generation", ylab = "Mean Kinship", 
               bg = "transparent");
          points(x = gens, y = mkval, pch = 20, cex = 1.5, type = "b",
                 lwd = 2, col = "black");
      }else{
          plot.new();
      }
  }else{
      x0 <- length(gens) - 5;
      x1 <- length(gens) + 0;
      par(mfrow = c(1, 2), mar = c(7, 5, 5, 5));
      hist(iter$kva, xlim = c(0, 1), ylim = c(0, 20), freq = FALSE, col = "black",
           yaxt = "n", ylab = "", main = "", xlab = "", 
           breaks = seq(from = 0, to = 1, by = 0.05), cex.axis = 1.5);
      title(xlab = "Kinship (k)", line = 4, cex.axis = 1.5, cex.lab = 1.5);
      box(lwd = 3);
      plot(x = x0:x1, y = 0:5, type = "n", ylim = c(0, 1), cex.axis = 1.5,
           cex.lab = 1.5, xlab = "Generation", ylab = "Mean Kinship", 
           bg = "transparent");
      points(x = gens, y = mkval, pch = 20, cex = 1.5, type = "b",
                 lwd = 2, col = "black");
  }
})
```

##

```{r lmSim2, fig.show='animate', echo = FALSE}
inputPanel(
  sliderInput("Imm2", label = "Immigrants", width="100%",
              min = 0.0, max = 10, value = 2, step = 1), 
  sliderInput("ID2", label = "Inbreeding Depression", width="100%",
              min = 0.0, max = 10, value = 8, step = 0.5), 
  sliderInput("cost2", label = "Strategy Cost", width="100%",
              min = 0.0, max = 0.1, value = 0, step = 0.01), 
  sliderInput("mu2", label = "Mutation Rate", width="100%",
              min = 0.0, max = 0.1, value = 0, step = 0.01), 
  actionButton("mate2", "New Generation"),
  actionButton("reset2", "Reset")
)

Ds    <- observeEvent(input$mate2, {
        if( is.null(gen$dat$ID) == TRUE | is.null(gen$dat$ID) == TRUE){
            gen$dat$ID   <- initialise_inds(N = 80);
            gen$dat$Rmat <- initialise_Rmat(gen$dat$ID);
        }else{
            gen$dat  <- new_gen(ID = gen$dat$ID, Rmat = gen$dat$Rmat, 
                                Kf = 40, Km = 40, beta = input$ID2, 
                                imm = input$Imm2, cost = input$cost2, 
                                mu = input$mu2);
            iter$all <- c(iter$all, mean(gen$dat$ID[,11:20]));
            iter$nue <- c(iter$nue, mean(gen$dat$ID[,21:30]));
            kvalues  <- gen$dat$Rmat;
            kvalues  <- kvalues[,-1];
            if(dim(kvalues)[1] > 80){
                kvalues <- kvalues[1:80, 1:80];
            }
            diag(kvalues) <- NA ;#diag(kvalues) * 0.5;
            iter$kva <- kvalues;
            iter$mkv <- c(iter$mkv, mean(kvalues, na.rm = TRUE));
        }
})

rs    <- observeEvent(input$reset2, {
       gen$dat <- list(ID = NULL, Rmat = NULL);
       iter$all <- 0;
       iter$nue <- 0;
       iter$kva <- 0;
       iter$mkv <- 0;
})

renderPlot({
  par(bg = 'transparent');
  malle <- iter$all;
  mneut <- iter$nue
  gens  <- 1:length(iter$all);
  if(length(gens) < 6){
      par(mfrow = c(1, 2));
      plot(x = 0:5, y = 0:5, type = "n", ylim = c(-1, 1), cex.axis = 1.5,
           cex.lab = 1.5, xlab = "Generation", ylab = "Inbreeding strategy", 
           bg = "transparent");
      abline(h = 0, lty = "dotted", lwd = 0.8);
      polygon(x=c(-0.125, 2.65, 2.65, -0.125), y=c(0.825, 0.825, 1.1, 1.1), 
              border=NA, col="black");
      polygon(x=c(-0.125, 2.65, 2.65, -0.125), y=c(-0.825, -0.825, -1.1, -1.1), 
              border=NA, col="black");
      abline(v = -0.125, lwd = 4);
      text(x = 0, y = 0.95, labels = "Inbreeding Preference", cex = 1.5, 
           col = "white", pos = 4);
      text(x = 0, y = -0.95, labels = "Inbreeding Avoidance", cex = 1.5, 
           col = "white", pos = 4);
      if(length(gens) > 0){
          points(x = gens, y = malle, pch = 20, cex = 1.5, type = "b",
                 lwd = 2, col = "black");
          par(mar = c(7, 5, 5, 5));
          hist(iter$kva, xlim = c(0, 1), ylim = c(0, 20), freq = FALSE, col = "black",
               yaxt = "n", ylab = "", main = "", xlab = "", 
               breaks = seq(from = 0, to = 1, by = 0.05), cex.axis = 1.5);
          title(xlab = "Kinship (k)", line = 4, cex.axis = 1.5, cex.lab = 1.5);
          box(lwd = 3);
      }
  }else{
      x0 <- length(gens) - 5;
      x1 <- length(gens) + 0;
      par(mfrow = c(1, 2));
      plot(x = x0:x1, y = 0:5, type = "n", ylim = c(-1, 1), cex.axis = 1.5,
           cex.lab = 1.5, xlab = "Generation", ylab = "Inbreeding strategy", 
           bg = "transparent");
      abline(h = 0, lty = "dotted", lwd = 0.8);
      polygon(x=c(x0-0.125, x0+2.65, x0+2.65, x0-0.125), 
              y=c(0.825, 0.825, 1.1, 1.1), border=NA, col="black");
      polygon(x=c(x0-0.125, x0+2.65, x0+2.65, x0-0.125), 
              y=c(-0.825, -0.825, -1.1, -1.1), border=NA, col="black");
      abline(v = x0-0.125, lwd = 4);
      text(x = x0, y = 0.95, labels = "Inbreeding Preference", cex = 1.5, 
           col = "white", pos = 4);
      text(x = x0, y = -0.95, labels = "Inbreeding Avoidance", cex = 1.5, 
           col = "white", pos = 4);
      points(x = gens, y = malle, pch = 20, cex = 1.5, type = "b",
                 lwd = 2, col = "black");
      par(mar = c(7, 5, 5, 5));
      hist(iter$kva, xlim = c(0, 1), ylim = c(0, 20), freq = FALSE, col = "black",
           yaxt = "n", ylab = "", main = "", xlab = "", 
           breaks = seq(from = 0, to = 1, by = 0.05), cex.axis = 1.5);
      title(xlab = "Kinship (k)", line = 4, cex.axis = 1.5, cex.lab = 1.5);
      box(lwd = 3);
  }
})
```

##

```{r, echo=FALSE}
OffATr <- function(r, m, mmin=1, Beta=1, c=1){
    return( (1/2)*(1+r) * (1 - exp(-c*(m - mmin - Beta*r)) ));
}
# Below is fmd for analysis and plotting
OffATrd <- function(r, m, mmin=1, Beta=1, c=1){
    return(  0.5 * (1 + r) * (exp(-c * (m - mmin - Beta * r)) * c)  );
    # Also can simplify: return(  (c/2)*(1+r)*exp(-c*(m-mmin-Beta*r))  );
} # We'll need the above to find m^{*} and \gamma^{*} 
# To find m^{*}, can use the function below
findm <- function(low.guess,high.guess,rval,mmin=1,Beta=1,c=1){
  fm  <- function(m,r=rval){ 
           OffATrd(r=rval, mmin=mmin, Beta=Beta, m=m, c=c)*(0-m) + 
             OffATr(r=rval, mmin=mmin, Beta=Beta, m=m, c=c);
         }
  lg  <- fm(m=low.guess, r=rval);
  hg  <- fm(m=high.guess,r=rval);
  if(lg > 0){
    u <- low.guess;
    l <- high.guess;
  }else{
    u <- high.guess;
    l <- low.guess;
  }
  if((fm(l) > 0 & fm(u) > 0) | (fm(l) < 0 & fm(u) < 0)){
    return("Value of m is outside the range");
  }else{
    check  <- 1;
    mguess <- 0.5 * (l+u);
    i      <- 0;
    while(abs(check) > 0.000001 & i < 1000000){
        check <- fm(r=rval, m=mguess);
        if(check > 0){
          u      <- mguess;
          mguess <- 0.5*(l+mguess); 
        }else{
          l      <- mguess;
          mguess <- 0.5*(u+mguess);
        }
        i <- i+1;
    }
    return(mguess);
  }
} # Running the below returns the estimate
fmpl  <- expression(zeta[off] == frac(1,2)*(1+r)*(1-e^{(-c*(m-m[min]-beta*r))}));
inputPanel(

  sliderInput("rval", label = "Relatedness", width="100%",
              min = 0.0, max = 1, value = 0, step = 0.05),

  sliderInput("cval", label = "PI curve", width="100%",
              min = 0.0, max = 3, value = 1, step=0.2),

  sliderInput("mmin", label = "Minimum m value", width="100%",
              min = 0.0, max = 2, value = 1, step = 0.1),

  sliderInput("Beta", label = "Inbreeding depression", 
              width="100%", min = 0.0, max = 5, value = 0, step = 0.2)
)

renderPlot({
  cc  <- as.numeric(input$cval);
  mm  <- as.numeric(input$mmin);
  bb  <- as.numeric(input$Beta);
  rr  <- as.numeric(input$rval);
  rmm <- findm(low.guess=0,high.guess=6,rval=rr,mmin=mm,Beta=bb,c=cc);
  mzt <- OffATr(r=rr, m=rmm, mmin=mm, Beta=bb, c=cc);
  rmg <- mzt / rmm;    
  PI  <- seq(from=0,to=4,by=0.01);
  All <- OffATr(r=rr, m=PI, mmin=mm, Beta=bb, c=cc);
  par(mar=c(5,5,2,2));
  plot(PI,All,type="l",lwd=4,ylim=c(0,1),
     xlab=expression(paste("Parental investment (",italic(m),")")),
     ylab=expression(paste("IBD alleles in offspring (",zeta[off],")")),
     cex.lab=1.5,cex.axis=1.5);
  abline(h=0,lty="dotted",lwd=0.8);
  abline(a=0,b=rmg,lty="solid", lwd=2,col="blue");
  arrows(x0=rmm,y0=mzt/2,x1=rmm,y1=0,length = 0.15,angle=30,code=2,lwd=3,col="blue");
  text(x=1,y=0.8,labels=fmpl,cex=2);
  text(x=rmm,y=0.5*mzt+0.07,labels=expression(paste(italic(m)^"*")),col="blue",cex=2);
  text(x=0.5,y=rmg*0.5+0.07,labels=expression(paste(italic(gamma^"*"))),
       srt=rmg,cex=2,col="blue");
})
```